\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A New Radial Time Projection Chamber for CLAS at JLab}

### 2017-01-17 eg6-letters/NIM/EG6_RTPC_NIM.tex
more updates to the NIM paper

\section{TPC technology} \label{prev_tpc}
In the late 1970s, David R. Nygren presented the concept of the Time projection 
Chambers (TPCs) as a new tool to track charged particles based on combining 
electric and magnetic fields within a sensitive gas volume. The ionized 
electrons are collected by a position-sensitive electron collection system.  
Later in 1997, F.~Sauli introduced the Gas Electron Multiplier (GEM) as an 
effective gas-gain element to improve the tracking efficiency of the TPCs 
\ref{gem_sauli}. Currently, the TPCs are widely used in physics experiments, 
such as high-energy physics (ALEPH, DELPHI), heavy-ion physics (STAR, ALICE) 
and rare-event searches (XENON, NEXT).

The first implementation of a radial TPC with JLab-CLAS spectrometer has been 
performed by the BoNuS experiment \cite{BONUS} in 2005. The success of the 
BoNuS-RTPC in detecting the slow protons motivated the development of this new 
RTPC. For the channel of interests of our experiments, the recoiled $^{4}He$ 
nuclei have momentum per charge around 150 MeV/c, the detection of these 
particles was the primary motivation for designing this detector.

\section{CLAS-EG6 RTPC design} \label{sec_design}
The design of CLAS-EG6 RTPC has been chosen according to physics and 
experimental considerations. At 6 GeV incident electron beam energy, the recoil 
$^{4}He$ nuclei, from the coherent DVCS channel, have an average momentum (per 
charge) around 125 MeV/c, while the "standard" CLAS detects charged particles 
with a threshold of 250 MeV/c. On the experimental site, The chosen design has 
to fit within the available space of a 5 Tesla Solenoid magnet that surrounds 
the target \cite{Hyon-suk}. Figure \ref{fig:RTPC2} shows an experimental view 
of the RTPC before inserting it into the solenoid.

The CLAS-EG6 RTPC is a 200~mm long with a 150~mm diameter time projection 
chamber. While a charged particle traverses a gas-sensitive volume, the ionized 
electrons are released and drift outward perpendicularly to the beam direction.  
These electrons are then amplified by three layers of GEMs and detected on 4.45 
by 5~mm pads. As for CLAS-BoNuS, the EG6-RTPC has two detection parts but 
unlike BoNuS they are part of a single chamber. Each part has independent GEM 
amplification systems that cover around 80\% of the azimuthal angle. Under the 
effect of the electromagnetic field, the ionization electrons drift to the 
readout system on the external shell with a curved trajectory as can be seen in 
the figure~\ref{fig:RTPC_1_4}.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of the RTPC taken on a plane perpendicular to the 
beam line, see text for description of the elements.} \label{fig:RTPC_1_4}
\end{figure}

As seen in figure~\ref{fig:RTPC_1_4}, the RTPC has the following substructure, 
from the beam axis to the exterior::\\
\begin{itemize}
   \item The target extends along the RTPC's central z-axis, with a diameter of 
      6 mm. It is enclosed in a 27-$\mu$m-thick Kapton wall.

   \item The first gas gap extends from 3 mm to 20 mm radial distance. It is 
   filled with $^{4}He$ gas at 1~atm. During run, this region is swarmed by 
   Moller electrons induced by the beam, such that filling this region with a 
   light gas like $^{4}He$ minimizes secondary interactions. This region is 
   surrounded by a 4 $\mu$m thick window made of aluminized Mylar and connected 
to ground.

   \item The gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME).  This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode.

   \item The drift region is filled with the Ne-DME gas mixture, it extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the central axis. The electric field in this region is around 
      500~V/cm (???) and perpendicular to the beam. Field cages are installed 
      at both ends of the detector in this region. Progressive voltage 
      depending on radial position is applied on these to maintain a uniform 
      electric field perpendicular to the beam line at the edges of the RTPC.

   \item The GEM system is composed of three GEMs located at radii, 60 mm, 63 
      mm and 66 mm, they amplify the electrons to make a measurable signal. In 
      this configuration, the first GEM layer represents the anode.

   \item The readout board has internal radius of 69~mm and collects the 
      charges, preamplifiers are plugged directly on its outer side and are 
      transmitting the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM amplification has been mainly chosen for this detector for their 
flexibility, allowing the curved amplification surface. Also GEM are known to 
create less sparks than in other gaseous detectors, which is a good thing for 
the detection of slow nuclei like $^4$He. GEMs are made from an insulator layer 
(here Kapton) sandwiched between two 5 $\mu$m copper layers. The mesh of each 
GEM layer is chemically etched with 50$\mu$m diameter holes in double-conical 
cross section shapes as can be seen in figure~\ref{fig:GEMs}. A potential 
difference of 400 V is applied between the two copper layers of each GEM 
creating a very strong electric field in the holes. Such strong fields lead to 
high ionization from initial electrons and therefore amplification of the 
signal. Also, a potential difference of 150 V is set between the GEMs to push 
the amplified electrons toward the readout pads. The gain of each GEM layer is 
of order 100.\\

>>>>

\section{TPC technology} \label{prev_tpc}
In the late 1970s, David R. Nygren presented the concept of the Time projection 
Chambers (TPCs) as a new tool to track charged particles based on combining 
electric and magnetic fields within a sensitive gas volume. The ionized 
electrons are collected by a position-sensitive electron collection system.  
Later in 1997, F.~Sauli introduced the Gas Electron Multiplier (GEM) as an 
effective gas-gain element to improve the tracking efficiency of the TPCs 
\ref{gem_sauli}. Nowadays, the TPCs are widely used in physics experiments, 
such as high-energy physics (ALEPH, DELPHI), heavy-ion physics (STAR, ALICE) 
and rare-event searches (XENON, NEXT).

The first implementation of a radial TPC with JLab-CLAS spectrometer has been 
performed by the BoNuS experiment \cite{BONUS} in 2005. The success of the 
BoNuS-RTPC in detecting the slow protons motivated the development of this new 
RTPC. For the channels of interest in CLAS-EG6 experiment, the recoiled 
$^{4}He$ nuclei have momentum per charge around 150 MeV/c. The detection of 
these low-energy nuclei was the primary motivation for designing this detector.

\section{CLAS-EG6 RTPC design} \label{sec_design}
The design of CLAS-EG6 RTPC has been chosen according to physics and 
experimental considerations. At 6 GeV incident electron beam energy, the recoil 
$^{4}He$ nuclei, from the coherent DVCS channel, have an average momentum (per 
charge) around 125 MeV/c, while the "standard" CLAS detects charged particles 
with a threshold of 250 MeV/c. On the experimental site, the chosen design has 
to fit within the available space inside a 5 Tesla Solenoid magnet that 
surrounds the target \cite{Hyon-suk}. Figure \ref{fig:RTPC2} shows an 
experimental view of the RTPC before being inserted into the solenoid.

The CLAS-EG6 RTPC is a 200~mm long with a 150~mm diameter TPC. While a charged 
particle traverses a gas-sensitive volume, the ionized electrons are released 
and drift, under the effect of an applied electromagnetic field, outward 
perpendicularly to the beam direction. These electrons are then amplified by 
three layers of GEMs and detected by a curved readout system on the external 
shell as can be seen in the figure~\ref{fig:RTPC_1_4}. As for CLAS-BoNuS, the 
EG6-RTPC has two detection halves but unlike BoNuS they are part of a single 
chamber. Each half has independent GEM amplification systems that cover around 
80\% of the azimuthal angle.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of CLAS-EG6 RTPC taken on a plane perpendicular to 
the beam line, see text for description of the elements.} 
\label{fig:RTPC_1_4}
\end{figure}

As seen in figure~\ref{fig:RTPC_1_4}, the CLAS-EG6 RTPC has the following 
substructure, from the beam line to the exterior:\\
\begin{itemize}
   \item The target extends along the RTPC's central z-axis, with a diameter of 
      6 mm. It is enclosed in a 27-$\mu$m-thick Kapton wall.

   \item The first gas gap extends from 3 mm to 20 mm radial distances. It is 
      filled with $^{4}He$ gas at 1~atm. During run, this region is swarmed by 
      M\o ller electrons induced by the beam, such that filling this region 
      with a light gas like $^{4}He$ minimizes secondary interactions. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and connected to ground.                                                                
   \item The gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode.

   \item The drift region is filled with the Ne-DME gas mixture, it extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is around 500~V/cm 
      (???) and perpendicular to the beam. Field cages are installed at both 
      ends of the detector in this region. Progressive voltage depending on 
      radial position is applied on these to maintain a uniform electric field 
      perpendicular to the beam line at the edges of the RTPC.

   \item The electron amplification system is composed of three GEMs located at 
      radii, 60 mm, 63 mm and 66 mm. In this configuration, the first GEM layer 
      represents the anode.

   \item The readout board has internal radius of 69~mm and collects the 
      charges. Preamplifiers are plugged directly on the outer side and are 
      transmitting the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM system has been mainly chosen for this detector for its flexibility, 
allowing the curved amplification surface. Also GEMs are known to create less 
sparks than in other gaseous detectors, which provides a cleaner detection of 
slow nuclei like $^4$He. GEMs are made from an insulator layer, Kapton in our 
case, sandwiched between two 5 $\mu$m copper layers. The mesh of each GEM layer 
is chemically etched with 50$\mu$m diameter holes in double-conical cross 
section shapes as can be seen in figure~\ref{fig:GEMs}. A potential difference 
of 400 V is applied between the two copper layers of each GEM creating a very 
strong electric field in the holes. Such strong fields lead to high ionization 
from initial electrons and therefore amplification of the signal. Also, a 
potential difference of 150 V is set between the GEMs to push the amplified 
electrons toward the readout pads. The gain of each GEM layer is of order 100, 
making a total gain of about $10^{6}$.\\
